\chapter{Implementation File Structure}
\label{sec:appendix-filestructure}

This appendix provides an implementation-oriented file map for the AI collaborator architecture described in the main chapters. The listing focuses on the modules that are directly relevant for invitation flow, chat runtime, orchestration, and browser automation.

\section{High-Level Repository Layout}
\begingroup
\footnotesize
\begin{verbatim}
optimization-platform/
├── Solver-Panel/
│   └── src/app/functions/internal-modeler/
│       ├── internal-modeler.component.ts
│       ├── internal-modeler.component.html
│       └── collaboration/
│           ├── collaboration-master-user.ts
│           ├── collaboration-user-builder.ts
│           └── collaboration.contracts.ts
├── Modeler/
│   ├── src/
│   │   ├── modeler.js
│   │   ├── mountAIAgent.jsx
│   │   └── ModalAIAgent.tsx
│   ├── components/
│   │   └── assistant-ui/
│   │       ├── thread.tsx
│   │       └── thread-list.tsx
│   └── lib/
│       └── chatApi.ts
└── AI-Collaborator/
    ├── puppeteer-server/
    │   ├── index.js
    │   └── puppet.js
    └── server/
        ├── src/agent.py
        ├── langgraph.json
        └── pyproject.toml
\end{verbatim}
\endgroup

\section{File Responsibilities}
\begin{itemize}
    \item \textbf{\texttt{internal-modeler.component.ts/.html}}: Adds and wires the ``Invite AI Agent'' action (\texttt{inviteAI()}) and collaboration link handoff from frontend to the automation bridge via \texttt{POST /puppet}.
    \item \textbf{\texttt{collaboration-master-user.ts}}: Provides the \texttt{writeUrl} used by \texttt{inviteAI()} to pass the writable collaboration link to the bridge.
    \item \textbf{\texttt{modeler.js}}: Hosts the modeler runtime and exposes the \texttt{window.Modeler} interface (\texttt{updateModelViewHTML}, \texttt{getHTML}, \texttt{validateHTML}) used by the Puppeteer automation layer.
    \item \textbf{\texttt{mountAIAgent.jsx}}: Mounts the embedded assistant-ui chat modal into the modeler container and binds the runtime callbacks.
    \item \textbf{\texttt{ModalAIAgent.tsx}}: Implements the chat modal component with thread lifecycle management and model-context injection per user turn.
    \item \textbf{\texttt{thread.tsx}}, \textbf{\texttt{thread-list.tsx}}: assistant-ui components responsible for conversation rendering, user input, and interrupt/resume controls.
    \item \textbf{\texttt{chatApi.ts}}: Implements the LangGraph SDK client operations (thread creation, state retrieval, and streamed run invocation) against the LangGraph runtime API.
    \item \textbf{\texttt{agent.py}}: Defines the LangGraph state graph, graph nodes (\texttt{llm\_node}, \texttt{validate\_html}, \texttt{apply\_html}, \texttt{review\_html}, \texttt{fix\_html}, \texttt{update\_html}), prompt contracts, and interrupt-based HITL flow. This is the sole Python entrypoint; the graph is registered via \texttt{langgraph.json}.
    \item \textbf{\texttt{langgraph.json}}: LangGraph deployment manifest that registers \texttt{agent.py:agent} as the compiled graph and points to the runtime environment file.
    \item \textbf{\texttt{index.js}}: Node/Express bridge server exposing the automation endpoints (\texttt{POST /puppet}, \texttt{POST /update-html}, \texttt{GET /get-html}, \texttt{POST /validate-html}).
    \item \textbf{\texttt{puppet.js}}: Puppeteer lifecycle management and direct modeler DOM read/write/validate execution via \texttt{window.Modeler} calls.
\end{itemize}

\section{Cross-Module Execution Path}
The operational call chain spans the listed modules as follows:
\begin{enumerate}
    \item Frontend invitation starts in \texttt{Solver-Panel}: \texttt{inviteAI()} retrieves the \texttt{writeUrl} from the master collaboration object and sends it to the bridge via \texttt{POST /puppet}.
    \item The bridge (\texttt{index.js}) delegates to \texttt{puppetInit(link)} in \texttt{puppet.js}, which launches a Puppeteer Chrome session, authenticates, and navigates to the collaborative workspace.
    \item Chat interaction is driven by the \texttt{Modeler} components: \texttt{ModalAIAgent.tsx} injects current model state into each user turn and forwards it to LangGraph through \texttt{chatApi.ts}.
    \item The LangGraph server executes the state graph defined in \texttt{agent.py}: LLM generation, optional repair, validation, and human-feedback interrupt/resume cycles.
    \item Tool calls within the graph invoke bridge endpoints (\texttt{GET /get-html}, \texttt{POST /validate-html}, \texttt{POST /update-html}).
    \item Puppeteer methods in \texttt{puppet.js} apply updates to the active modeler DOM via \texttt{window.Modeler.updateModelViewHTML(newHTML)}.
    \item Updated state propagates to all collaborators through the platform's existing collaboration synchronization channel.
\end{enumerate}

\chapter{API Endpoint Reference}
\label{sec:appendix-api-endpoints}

This appendix documents the API surfaces used by the implemented AI-collaboration pipeline. It includes the Node automation bridge endpoints and the LangGraph runtime operations used by the frontend.

\section{Node Automation Bridge (\texttt{AI-Collaborator/puppeteer-server})}
\label{sec:appendix-api-node-bridge}

\begin{itemize}
    \item \textbf{\texttt{GET /}} \\
    Payload: none \\
    Purpose: Health check endpoint used for basic service availability verification. \\
    Response: Confirmation text.

    \item \textbf{\texttt{POST /puppet}} \\
    Payload: \texttt{\{"link": <writeUrl>\}} \\
    Purpose: Initializes the Puppeteer-controlled Chrome session, authenticates the AI agent account, and navigates to the provided collaborative write URL. \\
    Response: Confirmation text.

    \item \textbf{\texttt{POST /update-html}} \\
    Payload: \texttt{\{"newHTML": <markup>\}} \\
    Purpose: Applies generated model markup to the active modeler instance in the browser session via \texttt{window.Modeler.updateModelViewHTML(newHTML)}. \\
    Response: \texttt{\{"success": true|false\}}.

    \item \textbf{\texttt{GET /get-html}} \\
    Payload: none \\
    Purpose: Retrieves the current model markup from the active browser session via \texttt{window.Modeler.getHTML()}. \\
    Response: Current HTML payload as JSON.

    \item \textbf{\texttt{POST /validate-html}} \\
    Payload: \texttt{\{"newHTML": <markup>\}} \\
    Purpose: Validates candidate markup against modeler constraints via \texttt{window.Modeler.validateHTML(newHTML)} inside the Puppeteer session. \\
    Response: Validation result (success flag and error details).
\end{itemize}

\section{LangGraph Runtime API Operations (Frontend Usage)}
\label{sec:appendix-api-langgraph}

The frontend uses the \texttt{@langchain/langgraph-sdk} client against the configured runtime URL (\texttt{http://127.0.0.1:2024}) through thread/run operations. The graph identifier used is the one registered in \texttt{langgraph.json} under the key \texttt{agent}.

\begin{itemize}
    \item \textbf{\texttt{threads.create()}} \\
    Input: none \\
    Purpose: Creates a new persistent thread and returns the thread identifier stored by the assistant-ui runtime for subsequent run calls.

    \item \textbf{\texttt{threads.getState(threadId)}} \\
    Input: \texttt{threadId} \\
    Purpose: Loads persisted message state and pending interrupt metadata, enabling chat state recovery across UI reloads.

    \item \textbf{\texttt{runs.stream(threadId, assistantId, \ldots)}} \\
    Input: \texttt{input.messages}, optional \texttt{command} (for interrupt resume), \texttt{streamMode: ['messages', 'updates', 'custom']} \\
    Purpose: Starts a streamed run for a thread and returns incremental model outputs and state updates rendered token-by-token in the chat UI.
\end{itemize}

\chapter{System Prompt Specification}
\label{sec:appendix-system-prompt}

This appendix records the system-prompt specification used by the AI orchestration service for optimization-model generation, as defined in \texttt{AI-Collaborator/server/src/agent.py} (\texttt{system\_prompt\_generate\_html}). Beyond role definition, this prompt encodes several reliability mechanisms. It includes strict structural and validation constraints (mandatory blocks, allowed child tags, and forbidden nesting patterns) so that generated outputs remain compatible with the modeler's expected schema.

It also prescribes a deterministic execution order for generation (identify entities, define data, then construct model blocks), which reduces ambiguity in complex prompts. In addition, the prompt design supports example-based in-context guidance (one-shot and few-shot style examples), enabling the model to align with desired input-output behavior without retraining. The \texttt{\{current\_html\}} placeholder is filled at runtime with the active workspace snapshot retrieved via \texttt{GET /get-html}.

\begingroup
\footnotesize
\begin{verbatim}
## Role and Objective
You are an expert Mathematical Programming Assistant. Your specific task is to
translate natural language problem descriptions into the "Markup Language for MIPs".

You must act as both a mathematical translator and a syntax specialist. You
will output valid XML-style markup adhering strictly to the grammar and
validation rules defined below.

## Core Language Rules
1. Hyphen Suffix: Because the platform uses HTML Custom Elements, every single
   tag MUST end with a hyphen (e.g., <model->, <con->, <var->, <each->).
2. Mandatory Model Blocks: The <model-> tag MUST contain EXACTLY ONE of each
   of the following tags: <variables->, <constraints->, and <objectives->.
3. Implicit Summation: You do not need to use the <add-> tag within objectives
   or constraints; child elements of <obj-> and <con-> are interpreted as a
   sum by default.
4. JavaScript Data: The <data-> section must contain a <script-> tag where
   sets and parameters are defined using standard JavaScript.

## Strict Grammar, Validations, and Allowed Attributes

### 1. Structural Tags
- <model->:
  - Must contain exactly one <objectives->, <constraints->, <variables->.
    May also contain <script->.
  - Attribute: src
- <data->: Must contain <each-> or <script->. (Attribute: src)
- <variables->:
  - Allowed children: <var->, <each->, <script->
  - VALIDATION CRITICAL: cannot contain <con-> or <obj->
- <constraints->:
  - Allowed children: <con->, <each->, <script->
  - VALIDATION CRITICAL: cannot contain <obj->
- <objectives->:
  - Allowed children: <obj->, <each->, <script->
  - VALIDATION CRITICAL: cannot contain <con->

### 2. Core MIP Elements
- <obj-> (Objective Function):
  - Allowed children: <var->, <add->, <mul->, <constant->, <each->, <script->
  - VALIDATION CRITICAL: cannot contain another <obj->
  - Attributes (in objectives): name, sense, weight, priority, absTol,
    relTol, index
  - Attributes (if inside a variable): name, coefficient
- <con-> (Constraints):
  - Allowed children: <var->, <add->, <mul->, <constant->, <each->, <script->
  - VALIDATION CRITICAL: cannot contain another <con->
  - Attributes (in constraints): name, lb, ub, index. Do NOT use <lhs> or
    <rhs> tags.
  - Attributes (if inside a variable): name, coefficient
- <var-> (Variables):
  - Allowed children (only if in variables): <obj->, <con->
  - Attributes (in variables): name, integer, lb, ub, index
  - Attributes (if inside constraint/objective): name, index, coefficient

### 3. Mathematical Statements
- <add-> and <mul->:
  - Allowed children: <var->, <constant->, <add->, <mul->, <each->, <script->
  - Attribute: coefficient
- <constant->:
  - No children
  - Attributes: name, value (float), index, coefficient

### 4. Feature Tags (Loops and Logic)
- <each->:
  - Attributes: family, item, index, coefficient, if
- <script->:
  - VALIDATION CRITICAL: cannot have 'modified' without 'src' or
    'data-definition'
  - Attributes: src, data-definition, modified

## Execution Instructions
0. If the user asks a conceptual/read-only question and does not request model
   creation/update, respond concisely in plain text without generating markup.
1. Identify sets, parameters, decision variables, objective function,
   constraints.
2. Draft the <data-> section with JavaScript arrays/variables in a <script->
   tag.
3. Construct the <model-> section ensuring all <var-> elements are defined
   in <variables-> before they are used in <objectives-> or <constraints->.
4. Check your work against the VALIDATION CRITICAL rules.
5. For create/update requests, return only final markup:
   <model-> ... </model-> (and <data-> ... </data-> when needed).
6. Do not include markdown code fences, explanations, tool calls, or extra
   prose when returning markup.
7. Preserve valid unchanged parts of the current model unless asked to modify.
8. Keep names and indices consistent across <data->, <variables->,
   <constraints->, and <objectives->.

## Example Output Format
  <model->
    <objectives->
      <obj- name="TotalCost" sense="minimize">
        <each- item="i" family="Items">
          <var- name="x" index="i" coefficient="cost[i]"></var->
        </each->
      </obj->
    </objectives->
    <constraints->
      <each- item="j" family="Bins">
        <con- name="Capacity" index="j" ub="capacity[j]">
          <each- item="i" family="Items">
            <var- name="x" index="[i,j]"></var->
          </each->
        </con->
      </each->
    </constraints->
    <variables->
      <each- item="i" family="Items">
        <each- item="j" family="Bins">
          <var- name="x" index="[i,j]" lb="0" ub="1" integer></var->
        </each->
      </each->
    </variables->
  </model->
  <data->
    <script->
      let Items = [1, 2, 3];
      let Bins = ['A', 'B'];
      let cost = [10, 20, 30];
      let capacity = {'A': 50, 'B': 50};
    </script->
  </data->

### Current Model

{current_html}
\end{verbatim}
\endgroup
