\chapter{Introduction}
\section{Context}
\subsection{Mixed-Integer Linear Programming}
Linear Programming (LP) \cite{wiki:Linear_programming} is a mathematical optimization framework in which a linear objective function is optimized under a set of linear equality and inequality constraints. From a methodological perspective, LP provides a rigorous way to represent decision problems in a formal, machine-solvable structure. Owing to this property, LP has become a foundational technique in operations research, applied mathematics, and computer science, with established use in practical domains such as production planning, logistics, energy management, and economic decision support \cite{kunwar2022introduction}.

In many real-world scenarios, however, decision variables are not purely continuous. Integer Linear Programming (ILP) \cite{wiki:Integer_programming} extends LP by constraining selected variables to integer domains, thereby enabling the representation of discrete alternatives, logical dependencies, and combinatorial decisions. Mixed-Integer Linear Programming (MILP) combines continuous and integer variables in a single formulation and is therefore particularly suitable for complex planning and scheduling settings in which continuous quantities (e.g., production volumes) and discrete decisions (e.g., on/off choices, assignment decisions, routing selections) must be optimized jointly. This expressive power explains why MILP is frequently used in both industrial and research contexts despite its increased modeling and computational complexity.

\subsection{Optimization Platform}
Despite their practical relevance, MILP models are often difficult to construct, validate, and maintain without substantial mathematical and technical expertise. Users must translate domain knowledge into formal constraints, ensure internal consistency of model components, and iteratively debug infeasible or suboptimal formulations. To reduce this barrier, prior work has developed a collaborative web-based Optimization Platform \cite{MarkoThesisMsc, DeniThesisMsc, MarkoThesisBsc, MaksimThesisMsc} that supports modeling and solving MILP instances through an interactive user interface.

The platform includes a visual drag-and-drop modeler, support for multiple parallel workspaces, and real-time collaboration between multiple users. Beyond usability, these features establish an iterative modeling workflow in which users can incrementally formulate, inspect, and refine optimization models while preserving shared context across collaborators. Consequently, the platform addresses both accessibility and productivity by lowering entry barriers for non-experts and reducing coordination overhead in team-based modeling scenarios.

\subsection{Human-AI Collaboration}
Recent advances in artificial intelligence, and in particular the emergence of transformer-based large language models (LLMs) \cite{vaswani2023attentionneed}, have enabled new forms of human-AI collaboration in technical workflows. Rather than functioning only as passive assistants, modern AI systems can participate in iterative problem solving, contextual interpretation, and tool-mediated task execution. In software and engineering contexts, this shift has already demonstrated that language-based interaction can reduce operational friction by allowing users to formulate complex intentions in natural language while delegating implementation-level actions to AI-supported systems.

Motivated by this development, this thesis investigates how AI-driven collaboration can be integrated into optimization modeling workflows to further improve platform usability, modeling efficiency, and support for users with heterogeneous levels of domain expertise. The central premise is that an AI agent can act not merely as a text-based helper but as an active collaborator embedded in the modeling environment, capable of proposing context-aware model modifications while remaining under explicit human supervision.

\section{Challenges}

Although the current optimization platform already provides a strong collaborative and user-oriented foundation, several limitations remain that constrain modeling efficiency, accessibility, and long-term usability in practice. In particular, the following challenges motivate the need for an AI-assisted collaboration layer in the proposed system:

\begin{enumerate}
    \item \textbf{Inefficiency in Complex Modeling Workflows:} The visual drag-and-drop interface improves usability for small to medium formulations, but can become increasingly time-intensive for large or highly structured MILP models. Repetitive modeling actions, frequent structural adjustments, and manual consistency checks introduce significant overhead. As model complexity grows, this interaction pattern can reduce development velocity and increase the risk of modeling errors that are difficult to detect without systematic support mechanisms.
    
    \item \textbf{Accessibility and Learning-Curve Barriers:} The platform offers a rich set of functionalities, but this feature depth can create a substantial onboarding burden for users with limited optimization background. New users must simultaneously understand domain concepts, interface mechanics, and modeling conventions before they can contribute effectively. Consequently, the initial learning curve remains steep, which may limit adoption and delay productive use in educational and applied settings.
    
    \item \textbf{Limited Real-Time Guidance for Individual Users:} While collaborative sessions with multiple human participants can provide peer support, users working independently often lack immediate assistance during model construction, debugging, or refinement. In such scenarios, unresolved uncertainty can interrupt workflow continuity and lead to trial-and-error interactions that are both slow and error-prone. A context-aware assistant capable of delivering on-demand guidance and actionable model suggestions can therefore improve both user confidence and overall modeling quality.
\end{enumerate}


\section{Goal}

The objective of this thesis is to design, implement, and analyze an AI agent that functions as a first-class collaborator within an existing collaborative optimization platform. In contrast to conventional assistant paradigms that remain external to the modeling process, the proposed agent is embedded directly into the shared workspace and participates through the same collaboration interface used by human users. Operationally, the agent receives natural-language instructions, interprets them with respect to the current model state, and generates candidate modifications to model structures, constraints, and related workspace artifacts.

A central design principle of this work is human-centered control and accountable collaboration. Consequently, agent-generated modifications are treated as provisional artifacts rather than immediately committed updates. Users can inspect proposed changes, approve or reject them, and provide iterative feedback that is incorporated in subsequent generation cycles. This interaction model establishes a human-in-the-loop workflow that combines automation with user supervision and preserves human authority over final model decisions.

To improve reliability, the architecture incorporates a validation stage between generation and commitment. This stage performs syntactic and structural checks on generated model artifacts and filters invalid, inconsistent, or hallucinated outputs before they affect the shared workspace. As a result, the system aims to maintain modeling correctness while still benefiting from rapid AI-assisted iteration.

At a research level, this thesis seeks to answer the following research questions:

\begin{itemize}
    \item \textbf{RQ1 : Can an integrated, stateful, and supervised agent architecture reduce modeling effort for routine and repetitive tasks?}
    \item \textbf{RQ2 : Can such an architecture lower entry barriers for users with limited optimization expertise? }
    \item \textbf{RQ3 : Can it improve the efficiency and quality of collaborative MILP model development?}
\end{itemize}

The overarching goal is therefore not only to introduce AI-generated assistance, but to operationalize it as a reliable collaboration mechanism within a production-oriented optimization environment.

\section{Solution Outline}

This section provides a structured overview of the proposed system architecture and its interaction workflow. At a high level, the solution combines four tightly coupled elements: (i) a collaborative modeling workspace, (ii) an LLM-based agent that interprets user intent, (iii) an iterative feedback mechanism for human supervision, and (iv) a validation layer that enforces model-level correctness constraints before commitment. Together, these components form a stateful human-AI collaboration loop designed for optimization modeling tasks.

\subsection{AI-Agent Collaborator}
The AI agent is integrated into the Optimization Modeler as a workspace participant rather than an external utility. After a user creates or opens a model, the agent can be invited through the existing collaboration infrastructure and becomes part of the shared session context. This design choice preserves architectural consistency with human collaboration while enabling the agent to access the same model state, interaction history, and relevant workspace metadata required for context-aware assistance.

Functionally, the agent acts as an execution-capable collaborator: it receives user instructions, maps them to model-level operations, and generates candidate changes in a form that can be inspected within the same collaborative environment. By embedding the agent directly into the session, the system avoids context fragmentation between separate tools and supports a unified modeling workflow.

\subsection{Iterative Feedback}
The collaboration workflow follows an iterative proposal-and-review pattern. Users express modeling intentions in natural language, after which the agent generates a candidate update based on the current workspace state. Instead of committing changes immediately, the system exposes the proposal for explicit user evaluation. Users can approve the proposal, reject it, or provide corrective feedback that is incorporated into subsequent iterations.

This interaction model serves two goals simultaneously: it leverages AI assistance to accelerate routine modeling operations, and it preserves human control over semantically critical decisions. Iterative refinement also enables progressive convergence toward higher-quality model formulations, especially in cases where the initial prompt is ambiguous or under-specified.

\subsection{Safety and Validation}
A dedicated validation stage is executed between agent generation and final model commitment. This stage performs syntactic and structural checks to ensure that generated artifacts are compatible with platform constraints and do not violate core modeling assumptions. Outputs that fail validation are rejected or returned for further refinement, preventing malformed or hallucinated updates from being applied to the shared workspace.

From a systems perspective, this validation layer is essential for reliable AI integration in optimization environments: it acts as a safeguard against invalid transformations while maintaining workflow continuity for valid proposals. Consequently, safety is treated as an operational property of the collaboration pipeline rather than a post hoc correction step.



