\chapter{Related Work}
\section{LLMs and Mathematical Modeling}
\subsection{End-to-End Modeler and Solver Archirecture}

Extensive research has been conducted at the intersection of Large Language Models (LLMs) and mathematical optimization. A prominent example is \textbf{OptLLM} \cite{zhang2024solvinggeneralnaturallanguagedescriptionoptimization}, a framework designed to bridge natural language descriptions and specialized solvers. OptLLM utilizes a modular architecture consisting of three primary components: the \emph{Interaction Refinement Module}, which processes user queries and handles multi-round clarifications; the \emph{Converter Module}, which includes a formulator to translate text into mathematical structures and a coder to generate executable code; and finally, the \emph{Responser Module}, which executes the model via external solvers and interprets the results for the user. This architecture demonstrates the feasibility of integrating disparate modules to transform an LLM into a comprehensive end-to-end optimization system. 

However, OptLLM operates as a standalone, sequential pipeline: it does not embed the agent into a shared modeling environment, does not support multi-user collaborative workspaces, and provides no human-in-the-loop mechanism for reviewing or rejecting individual model updates before they are committed. The architecture proposed in this thesis addresses these gaps by integrating the AI agent directly into the live collaborative workspace as a first-class participant, and by wrapping every proposed change in an explicit approval cycle that keeps the human modeler in authority throughout the editing process.


\subsection{Natural Langauge to Formal Specification}
Another significant contribution is \textbf{LLMOPT} \cite{jiang2025llmoptlearningdefinesolve}, a learning-based framework that focuses on boosting the generalization of LLMs across various optimization types. LLMOPT introduces a universal \emph{Five-Element Formulation} that reduces complex problem descriptions into five core components: sets, parameters, variables, objectives, and constraints. This structured representation allows the system to generate consistent solver code across diverse domains, including linear, nonlinear, and combinatorial optimization.

\subsection{LLM Benchmark for Optimization Modeling}
While the aforementioned frameworks focus on generation, \textbf{Mamo} \cite{huang2025llmsmathematicalmodelingbridging} provides a specialized benchmark to evaluate these modeling capabilities. Mamo consists of 1,209 meticulously curated questions spanning Ordinary Differential Equations (ODEs), Linear Programming (LP), and Mixed-Integer Linear Programming (MILP). Unlike traditional result-oriented metrics, Mamo utilizes a ``model generation and numerical validation'' approach. An LLM is tasked with formulating a model from the dataset, which is then executed by a solver; the resulting numerical answer is compared against the provided ground truth. 

Complementary to Mamo, the \textbf{NLP4LP} dataset \cite{ahmaditeshnizi2023optimusoptimizationmodelingusing} was introduced alongside the \textbf{OptiMUS} framework to further benchmark the formulation of LP and MILP problems from natural language descriptions. While OptiMUS serves as an autonomous LLM-based agent capable of drafting, debugging, and testing solver code, its accompanying NLP4LP dataset provides a rigorous testing ground for evaluating such capabilities. By challenging agents to generate executable mathematical models from text and verifying the resulting solutions, NLP4LP establishes a clear baseline for measuring how effectively an AI can interpret complex optimization constraints. Together, datasets like Mamo and NLP4LP provide the necessary quantitative foundation for assessing the accuracy and robustness of LLM-driven modeling tools. Crucially, a selected subset of the NLP4LP dataset serves as the core benchmark for the experimental evaluation conducted later in this thesis.


\section{AI Code Generation}

\subsection{Current State}
Since the introduction of the Transformer architecture \cite{vaswani2023attentionneed}, there has been a paradigm shift in the field of automated software engineering. While Transformers were initially celebrated for their success in Natural Language Processing (NLP), their ability to model long-range dependencies and complex structural patterns has led to significant advancements in Large Language Models (LLMs) for code generation.

Extensive research has explored the transition from general-purpose text generation to domain-specific code synthesis. As detailed in the comprehensive survey by Jiang et al. \cite{Jiang_2026}, modern frameworks leverage massive datasets of source code to perform tasks such as code completion, bug detection, and natural-language-to-code translation. The survey categorizes state-of-the-art models, including GPT-4, CodeLlama, and StarCode, and examines the underlying training objectives, such as ``Fill-in-the-Middle'' (FIM) and instruction tuning, which are essential for understanding the non-linear structure of programming languages.

\subsection{Coding Agents}

Recent advancements in AI-powered coding assistants such as GitHub Copilot \cite{copilot}, Google's Gemini CLI \cite{geminicli}, and Anthropic's Claude Code \cite{claudeCodeAgent} demonstrate how generative AI significantly enhances software development workflows. These coding agents operate as collaborative tools that integrate directly into development environments, where they utilize Large Language Models (LLMs) fine-tuned on vast code and documentation corpora \cite{yang2025codefoundationmodelsagents}. Typically, users write natural language prompts or partial code snippets; the agent responds by generating, refactoring, or completing code, suggesting best practices, and occasionally identifying bugs.

Most modern coding agents rely on transformer-based architectures pre-trained on publicly available open-source repositories, and are further tuned for developer intent and task relevance through techniques such as Reinforcement Learning from Human Feedback (RLHF) and tool integration APIs \cite{yang2025codefoundationmodelsagents}. For example, GitHub Copilot, powered by OpenAI's Codex model, integrates seamlessly into IDEs and proactively suggests entire code blocks in real-time. Google's Gemini CLI and Anthropic's Claude Code Agent similarly offer advanced contextual understanding, enabling cross-file reasoning and multi-modal support such as code explanation, completion, or test synthesis.

Despite their power, these tools exhibit key limitations when applied to mathematical optimization modeling \cite{forootani2025surveymathematicalreasoningoptimization}. Primarily, their training prioritizes general-purpose software engineering tasks, such as web development, scripting, or library usage, rather than symbolic reasoning or formal specification needed for optimization problems. As a result, these agents may struggle with generating accurate and semantically valid mathematical models, including algebraic constraints, variable declarations, and objective functions, particularly when such constructs must strictly adhere to domain-specific formats. Furthermore, their validation mechanisms are largely confined to code execution or basic unit tests, whereas optimization modeling requires both syntactic correctness and adherence to deeper mathematical requirements.

Thus, while coding agents have greatly improved code generation and developer productivity in mainstream programming, their architecture and underlying datasets are not inherently optimized for the highly structured, domain-specific tasks characteristic of mathematical optimization modeling.


\subsection{LLM Benchmarking for Code Generation}
As Large Language Models (LLMs) continue to evolve, the demand for more rigorous evaluation methodologies has led to the development of advanced benchmarking frameworks. One such framework is \textbf{EvalPlus} \cite{liu2023codegeneratedchatgptreally}, which addresses the limitations of traditional code generation benchmarks. 

EvalPlus builds upon established datasets, most notably \emph{HumanEval} \cite{chen2021evaluatinglargelanguagemodels} and \emph{MBPP}. While these original benchmarks provided a foundational method for measuring functional correctness, their test suites are often sparse and fail to catch edge-case failures, allowing incorrect code to pass as correct. EvalPlus mitigates this by automating the synthesis of thousands of additional test cases for each problem. By significantly increasing the test-case density, often by a factor of 80x compared to the original suites, EvalPlus creates a more robust evaluation environment that exposes ``false positives'' in LLM-generated code. 


\section{Human-in-the-loop Architecture}

\subsection{Agentic Development Environments}

Recent years have witnessed the emergence of agentic Integrated Development Environments (IDEs) that proactively embed AI agents into the user’s workflow. Notable examples include \textbf{Cursor} \cite{cursor}, Google’s \textbf{Antigravity} \cite{antigravity}, and the Copilot extensions for Visual Studio Code \cite{visualstudio}. These environments aim to facilitate a tight human-in-the-loop collaboration by making intelligent assistants easily available during the development process.

For instance, \textbf{Cursor} offers an AI agent that not only generates and edits code but also explains suggestions, processes natural language commands, and supports high-level refactoring through conversational interfaces directly within the code editor. Google's \textbf{Antigravity} further pushes this concept by providing contextual tool recommendations and automated fixes based on the developer’s current task, often prompting the user for confirmation or guidance before performing significant actions. Similarly, Copilot in Visual Studio Code integrates an AI agent trained on vast open-source codebases, providing real-time code completions and suggestions that the human developer can freely accept, reject, or modify as needed.

In all these platforms, the human developer remains a principal decision-maker reviewing, editing, or vetoing AI-generated proposals. This forms a continuous feedback loop where the agent assists, but all critical actions are subject to user oversight and approval, embodying the essence of ‘‘human-in-the-loop’’ design for code authoring environments.

However, while these agentic IDEs excel in integrating AI assistance for general-purpose software development, their capabilities are typically confined to code generation, editing, and debugging. They are not inherently equipped for specialized tasks such as mathematical modeling, formal specification, or symbolic reasoning outside traditional software engineering domains. As a result, their human-in-the-loop mechanisms and agentic workflows remain primarily limited to mainstream coding activities, rather than broader, domain-specific applications like mathematical optimization modeling.


\subsection{Domain-Specific Agentic Systems}

Recent advancements have witnessed the emergence of domain-specific agentic systems, AI assistants designed with a focus on particular industries, products, or workflows, rather than general-purpose programming. These agents are embedded into platforms to automate, streamline, and augment specialized tasks through close integration, fine-tuned models, and continual user-driven feedback.

One notable example is \textbf{v0} \cite{v0}, an AI-powered design assistant specializing in rapid user interface (UI) prototyping. V0 enables designers to describe layouts and components in natural language, then instantly generates production-grade code tailored to target frameworks like React or Vue. The system’s effectiveness is a result of fine-tuning large language models on datasets encompassing real-world design systems, annotated UIs, and established coding conventions, ensuring that outputs are compatible with prevailing design and engineering standards.

Industry-leading productivity suites have also embraced agentic AI. Both the \textbf{Microsoft Copilot suite} \cite{microsoftCopilot} (integrated into Word, Excel, PowerPoint, Outlook, etc.) and the AI features within the \textbf{Google Workspace} \cite{googleWorkspace} suite (Docs, Sheets, Slides, Gmail, and more) exemplify domain-specific assistants for business documentation and office workflows. These agents draft emails, generate meeting summaries, synthesize data into insights, automate spreadsheet analyses, and manage scheduling, all by leveraging AI models refined with proprietary enterprise data, fine-tuned user interactions, and organization-specific context. Their integration is further enhanced by enforcing strict access controls, supporting privacy-preserving computations, and continuously learning from authenticated user behavior within organizational environments.

In the academic and technical writing domain, platforms like \textbf{Overleaf} \cite{overleaf} have integrated targeted AI helpers that support LaTeX authors by providing syntax correction, auto-completion of citations, formatting suggestions for mathematical content, and real-time collaborative editing tips. These capabilities are achieved by specializing large language models with academic writing data and technical document corpora, ensuring that suggestions are both technically accurate and contextually appropriate.

In addition to these deep integrations, a ubiquitous trend is the proliferation of AI helper chatbots directly on product, SaaS, and documentation pages. A representative example is the AI assistant available on the \textbf{Assistant-UI} documentation page \cite{assistantui}, where an integrated chatbot guides users through setup processes, answers technical questions, and helps to resolve onboarding or configuration issues in real time. Such conversational agents are designed to offer immediate, context-aware assistance, addressing product queries, helping new users get started, and providing proactive troubleshooting or feature recommendations. Their models are typically trained on product documentation, frequently asked questions, and anonymized user interactions, ensuring that responses are both relevant and tailored to the specific service. Many vendors fine-tune these models to reflect evolving user needs and workflows, continually enhancing the effectiveness and relevance of automated customer support.

Collectively, these domain-specific agentic systems illustrate a shift toward AI agents that are purpose-built, data-driven, and closely aligned with the specialized tasks and environments of their target users. Through model fine-tuning, deep platform integration, and continual feedback, they offer contextually intelligent support that goes well beyond what is achievable with general-purpose language models.


\subsection{Magentic-UI}

\textbf{Magentic-UI} \cite{mozannar2025magenticuihumanintheloopagenticsystems} is a human-in-the-loop (HITL) research prototype by Microsoft Research, designed to facilitate complex web-based tasks through close collaboration between the user and an AI agent. In contrast to fully autonomous agents, Magentic-UI foregrounds the user's role as an active collaborator. The system processes natural language queries from the user, interprets intent, and generates a structured, multi-step plan for task execution.

Central to Magentic-UI are its advanced interaction mechanisms: \emph{co-planning} and \emph{action guards}. Through co-planning, the agent transparently presents its proposed sequence of actions, allowing the user to review, modify, approve, or halt steps before anything is carried out. Action guards further require explicit human approval for any pivotal or potentially sensitive operations, such as sharing confidential data or triggering transactions. This layered approach ensures user oversight at key decision points, fostering transparency and safety throughout the process. By embedding human review and iterative feedback into the agent's workflow, Magentic-UI exemplifies contemporary best practices in human-in-the-loop AI architectures, especially for web-based, agentic environments, highlighting the synergy of AI assistance and maintained human authority.

% \section{Schema Validation}