\chapter{Implementation}
This chapter translates the conceptual architecture into the implemented system. It describes how the AI collaborator is integrated into the existing Optimization Platform, how requests move across services, and how generated updates are applied to shared workspaces in a human-in-the-loop workflow. For readers who are new to the platform internals, it is recommended to first review the prior thesis work that established the base collaboration and modeler architecture \cite{MarkoThesisMsc, DeniThesisMsc, MarkoThesisBsc, MaksimThesisMsc}. Doing so provides the full system context and makes the implementation extensions introduced in this chapter easier to follow.

\section{Implementation Overview}
\subsection{System Components}
The implementation is organized around four concrete components:
\begin{enumerate}
    \item \textbf{Solver-Panel (Angular):} User-facing entry point for authentication, collaboration, and modeler access.
    \item \textbf{Modeler runtime (JavaScript/React integration):} Hosts the visual modeling canvas and embeds the AI chat interface.
    \item \textbf{Agent orchestration service (Python):} Executes stateful LLM workflow logic and human-feedback loops using LangGraph \cite{langchainLangGraph}.
    \item \textbf{Automation bridge (Node.js):} Controls browser-level interactions and DOM-side workspace updates through Puppeteer \cite{pptrPuppeteerPuppeteer, chromiumHome}.
\end{enumerate}

This decomposition preserves existing platform functionality \cite{MarkoThesisMsc, DeniThesisMsc, MarkoThesisBsc, MaksimThesisMsc} while adding AI collaboration as an orthogonal capability.

\subsection{End-to-End Request Path}
At runtime, the high-level request path is:
\begin{enumerate}
    \item A user invites the AI collaborator into an active editable workspace.
    \item The frontend mounts and exposes a chat interface inside the modeler context.
    \item User messages are sent through the chat runtime to the orchestration backend.
    \item The orchestration service builds context from conversation state and current model state.
    \item If required, tool calls trigger browser-layer operations through HTTP endpoints.
    \item The resulting model update is applied to the shared workspace and becomes visible to collaborators.
    \item The user may accept, reject, or refine through an interrupt-driven feedback loop.
\end{enumerate}

\subsection{Relevant Files and Project Structure}
To make the implementation traceable, this section highlights the primary code locations used in the AI-collaboration pipeline:
\begin{itemize}
    \item \textbf{Solver-Panel integration:} internal modeler component files where invitation and collaboration handoff are implemented.
    \item \textbf{Modeler chat runtime:} modeler files responsible for mounting the assistant UI, injecting model context, and forwarding thread/run calls.
    \item \textbf{Python orchestration:} LangGraph workflow definition, tool binding, and prompt-control logic.
    \item \textbf{Node automation bridge:} HTTP endpoints and Puppeteer methods for browser session initialization and model DOM read/write operations.
\end{itemize}

Because the repository is a multi-module workspace, a complete structure listing is provided in the appendix for reproducibility and navigation. The detailed tree and file-role mapping are documented in \cref{sec:appendix-filestructure}.

\section{Frontend Integration}
For detailed API endpoint documentation related to this section (invitation, runtime communication, and service calls), see \cref{sec:appendix-api-endpoints}.

\subsection{Agent Invitation}
The integration extends the existing collaboration workflow by adding an ``Invite AI Agent'' action in the internal modeler view, implemented through the \texttt{inviteAI()} method in \texttt{internal-modeler.component.ts}. This method is the concrete frontend entry point for agent invocation.

Operationally, \texttt{inviteAI()} performs the following sequence:
\begin{enumerate}
    \item Ensure that a collaboration context exists; if not, create one through the existing master-user collaboration builder.
    \item Retrieve the workspace edit link from the master collaboration object (the write-capable collaboration URL).
    \item Send an HTTP request to the automation bridge endpoint:
    \[
    \texttt{POST /puppet}
    \]
    with JSON payload:
    \[
    \{\texttt{"link"}:\ \texttt{<writeUrl>}\}
    \]
    \item Update frontend UI state to expose the AI collaborator controls after successful invocation.
\end{enumerate}

On the backend side, this endpoint is implemented in the Node/Express bridge service (\texttt{AI-Collaborator/puppeteer-server/index.js}) as \texttt{POST /puppet}. When called, the endpoint invokes a Puppeteer initialization routine that launches a server-side Chrome session. The automation routine logs the agent into the platform using dedicated credentials, then navigates to the provided collaborative write URL. This browser session serves as the AI agent's execution environment, enabling the agent to join the current collaboration session as an editing participant and interact with the same live modeler state as human users.

The invitation path remains role-aware because it reuses the platform's existing edit/read collaboration semantics. Therefore, no parallel permission model is introduced; the AI participant is integrated through the same collaboration boundary already used for human users.

\subsection{Embedded Chat Runtime}
The chat interface is mounted directly in the modeler container, not as an external application. The implementation uses assistant-ui with a React-based modal/thread runtime \cite{assistantui}, so user interaction remains co-located with model editing.

Communication between assistant-ui and LangGraph is implemented through the LangGraph runtime adapter in the frontend. Concretely, the runtime defines three core operations: \texttt{create}, \texttt{load}, and \texttt{stream}. During \texttt{create}, the frontend requests a new thread from the LangGraph API and stores the returned external thread identifier in the assistant-ui runtime state. During \texttt{load}, the frontend retrieves the thread state from LangGraph and reconstructs the local conversation view (messages and pending interrupts). This ensures that chat state remains persistent across UI reloads and modal reopen events.

During active interaction, assistant-ui invokes the \texttt{stream} operation for each user turn. The frontend sends the message payload (plus runtime command metadata) to the LangGraph run-stream endpoint and subscribes to incremental response events. These streamed events are rendered token-by-token in the assistant-ui thread, which provides immediate user feedback and preserves conversational continuity. In parallel, command responses from LangGraph (e.g., interrupt/resume requirements in the human-in-the-loop flow) are surfaced in the same thread UI so that users can explicitly approve, reject, or refine the current proposal.

At the protocol level, assistant-ui is therefore responsible for interaction orchestration in the browser (thread lifecycle, rendering, user actions), while LangGraph is responsible for stateful agent execution on the backend (graph transitions, tool calls, and interrupt logic). The integration boundary is a thread-oriented API contract: assistant-ui sends user turns and commands, and LangGraph returns streamed model outputs plus state updates.

\subsection{Model Context Injection}
For each conversational turn, the frontend retrieves the current state of the active modeler instance and injects it into the request context before forwarding the user message. This mechanism is implemented in the chat runtime path by querying the modeler state (HTML/instance representation) and prepending it as a system-level context message for the agent run.

The injected context serves two implementation goals. First, it grounds the model in the \emph{current} workspace snapshot, including already defined variables, constraints, objectives, and structural nesting. Second, it provides enough local state to interpret referential instructions such as ``modify the existing capacity constraint,'' ``add this variable to the objective,'' or ``update the bound we added earlier.'' Without explicit context injection, such instructions would be ambiguous or incorrectly resolved.

From a data-flow perspective, each user turn is transformed into a composite payload:
\begin{enumerate}
    \item current workspace/model representation (system context),
    \item conversational history (thread messages),
    \item current user instruction (new turn),
    \item optional runtime command metadata (e.g., interrupt resume input).
\end{enumerate}

This design improves consistency across iterative runs, because the agent does not rely on implicit memory alone; it receives an explicit state snapshot at invocation time. In practice, this reduces context drift and helps keep generated updates aligned with the latest model state, which is particularly important in collaborative sessions where the workspace can change between turns.

\section{Agent Orchestration Service}
\subsection{LangGraph Workflow Realization}
The orchestration backend is implemented as a state graph with explicit nodes and conditional transitions \cite{langchainLangGraph}. In API terms, the frontend and backend communicate through LangGraph's thread-and-run abstraction introduced in the embedded chat subsection:
\begin{itemize}
    \item \textbf{Thread:} persistent conversation/state container for one chat session.
    \item \textbf{create:} initializes a new thread and returns an external thread identifier.
    \item \textbf{load:} retrieves stored thread state (messages and pending interrupts).
    \item \textbf{stream:} starts a run on a thread and returns incremental output events.
    \item \textbf{command:} runtime control input, used for actions such as interrupt resume.
\end{itemize}

Operationally, assistant-ui uses these API primitives to drive LangGraph execution: each user turn is sent through a streamed run bound to a thread ID; LangGraph then executes the state graph, emits partial outputs, and persists updated state back to that same thread.

At graph level, the implemented flow includes:
\begin{itemize}
    \item an LLM node for proposal generation,
    \item validation and repair nodes that gate model updates,
    \item an apply node that writes changes to the live workspace,
    \item and a human-feedback node for interrupt/resume control.
\end{itemize}

This graph-level structure directly realizes the intended HITL execution loop:
\[
\text{request} \rightarrow \text{generation} \rightarrow \text{tool/action} \rightarrow \text{feedback} \rightarrow (\text{regenerate or terminate})
\]

More concretely, LangGraph executes this loop by evaluating conditional edges after each node execution. If the LLM response requires a tool call, control moves to the tool node; if human approval is required, an interrupt state is emitted and exposed to the frontend. The frontend then sends a command (resume payload), and graph execution continues from the interrupt point instead of restarting from scratch. This resume semantics is the key mechanism that turns a normal chat pipeline into a stateful human-in-the-loop workflow.

Compared to ad-hoc request handlers, this implementation provides explicit state transitions, observable control flow, deterministic routing conditions, and durable conversational state across turns.

\subsection{Graph States}
The implemented orchestration can be described as a set of explicit graph states, where each state has a narrowly defined responsibility in the request lifecycle.

\textbf{LLM State (Generation State).}
This state is responsible for proposal generation. It receives the composite context (thread history, current model snapshot, and current user instruction), applies system-level constraints, and produces either:
\begin{itemize}
    \item a candidate model update, or
    \item a tool-call request when external environment interaction is required.
\end{itemize}
In other words, this state determines \emph{what} should happen next.

\textbf{Execution State (Validation/Apply).}
These states execute external actions through HTTP calls to the automation bridge, such as validating candidate markup, reading current model content, or applying updates to the active workspace. The results of these calls are fed back into graph state so the next LLM step can reason over fresh environment results. This state therefore implements \emph{environment-grounded execution} rather than pure text generation.

\textbf{HITL State (Approval/Feedback State).}
This state materializes human-in-the-loop control through interrupt and resume semantics. When user confirmation is required, the graph emits an interrupt that is surfaced to the frontend. The user can then approve, reject, or provide refinement feedback; the returned command resumes graph execution from the suspended state. This keeps user authority inside the orchestration loop rather than outside it.

\textbf{Validation State (Reliability State).}
A dedicated validation state has been integrated into the implemented graph. This state executes before the final commit and verifies that AI-generated output conforms to modeler constraints and schema expectations. Instead of introducing an ad-hoc validator, the implementation reuses the validation logic already present in the modeler stack (i.e., existing modeler validators and checks) by calling the automation bridge \texttt{POST /validate-html} endpoint. This provides a deterministic acceptance gate:
\[
\text{generated output} \rightarrow \text{modeler validator checks} \rightarrow (\text{accept} \mid \text{return errors for refinement})
\]
and would further reduce the risk of structurally valid but semantically inconsistent updates.


\subsection{External Actions and Bridge Calls}
The orchestration service triggers external actions by calling the Node automation bridge over HTTP. Instead of a separate tool node, these actions are represented as explicit graph states that call bridge endpoints and feed responses back into the LangGraph state.

The relevant runtime actions are:
\begin{itemize}
    \item \textbf{\texttt{get\_current\_html()}}: Retrieves the current model representation from the active browser-side modeler instance via \texttt{GET /get-html}. This is used when the agent needs fresh environment state before generating or refining an update.
    \item \textbf{\texttt{validate\_html(new\_html)}}: Validates candidate markup using the modeler validators exposed through \texttt{POST /validate-html}. Validation errors are routed into the repair path.
    \item \textbf{\texttt{update\_html(new\_html)}}: Applies generated model content back to the live modeler DOM through \texttt{POST /update-html}. This is the execution step that materializes the agent proposal in the shared workspace.
\end{itemize}

In addition, development/testing helper tools (e.g., arithmetic helpers used during graph experiments) can be bound in the same mechanism, but they are not part of the production modeling flow.

These tools call the Node automation bridge via HTTP, which keeps environment execution separate from LLM reasoning logic.

This separation is architecturally important: the Python service decides \emph{what} change to perform, while the automation bridge performs \emph{how} the change is enacted in the browser session.

\subsection{Prompt and Constraint Strategy}
The implemented prompting strategy uses a strong \emph{system prompt} as a behavioral contract for the agent. In this context, the system prompt is the highest-priority instruction layer passed to the model at run time; it defines the agent role, task boundaries, expected output format, and non-negotiable constraints. User messages are interpreted within this envelope rather than in isolation.
For the full prompt specification used in this thesis implementation, see \cref{sec:appendix-system-prompt}.

For this thesis, the system prompt is designed around the optimization modeler's domain-specific markup language. It explicitly specifies:
\begin{itemize}
    \item required structural organization (root tags, mandatory blocks, nesting rules),
    \item allowed elements and attributes for variables, objectives, constraints, loops, and script blocks,
    \item and validation-critical prohibitions (forbidden nesting patterns and structurally invalid combinations).
\end{itemize}
This design reduces under-specified generation and keeps outputs aligned with the modeler's parser and execution expectations.

The prompt also helps in three practical ways. First, it improves \textbf{format reliability} by reducing malformed outputs. Second, it improves \textbf{semantic alignment} by reminding the model that it acts as a mathematical modeling assistant, not a generic chatbot. Third, it improves \textbf{repairability}: when validation errors occur, the same constraint framework can be reused to guide the model toward corrected output in the next iteration.

In addition to core instructions, the implementation can include \emph{example-based prompting}. Terminologically, this is often referred to as \textbf{one-shot} (one example) or \textbf{few-shot} (multiple examples) prompting. This is not model training in the machine-learning sense; instead, it is in-context conditioning at inference time. The examples demonstrate target input-output behavior (e.g., natural-language request to valid markup transformation), which helps the model imitate the desired structure and reduce ambiguity in difficult cases.

Overall, the prompt layer acts as a soft specification interface between natural-language intent and executable model transformations, complementing downstream validation and human-in-the-loop supervision.

\subsection{Validation Gate Before Model Application}
Before any AI-generated optimization model update is committed to the collaborative workspace, it should pass a validation gate based on the modeler package's existing validators/checks. The purpose of this gate is to ensure that generated output is not only well-formed text, but structurally and semantically compatible with the modeler's expected representation.

Conceptually, the validation pipeline is:
\[
\text{LLM output} \rightarrow \text{modeler validators} \rightarrow (\text{accept and apply} \mid \text{reject and refine})
\]

In this design, validation is executed before DOM mutation. If the validator layer accepts the generated artifact, the update is forwarded to the automation execution path and applied to the active modeler instance. If validation fails, the error diagnostics are returned to the orchestration loop and used as structured feedback for the next generation attempt. This keeps invalid or partially inconsistent updates from propagating to collaborators.

Reusing modeler-native validators has two implementation advantages. First, it aligns AI output quality checks with the same rule set already used by the platform, avoiding divergence between AI and manual editing paths. Second, it creates a deterministic acceptance criterion that can be audited and improved incrementally as the validator suite evolves.

\section{Browser Automation Execution Layer}
\subsection{Session Initialization}
The Node.js automation bridge exposes HTTP endpoints for browser lifecycle actions. Initialization starts when the frontend invitation flow calls:
\[
\texttt{POST /puppet}
\]
with payload containing the collaborative write URL. In the bridge service, this endpoint delegates to the Puppeteer bootstrap method (\texttt{puppetInit(link)}), which performs the complete startup sequence.

Concretely, initialization executes the following steps:
\begin{enumerate}
    \item \textbf{Browser launch:} start a Puppeteer-controlled Chrome instance on the server side and create a new page context.
    \item \textbf{Viewport/session setup:} configure runtime viewport and browser page state used for subsequent modeler interaction.
    \item \textbf{Authentication flow:} navigate to the platform login route and submit AI-account credentials through scripted form interactions.
    \item \textbf{Workspace attach:} after successful login, navigate to the received collaboration write URL.
    \item \textbf{Persistent page binding:} store the initialized page handle for later read/write calls (model retrieval and model update endpoints).
\end{enumerate}

After this sequence, the automation layer is attached to the same live collaborative modeler session as human users. Subsequent tool calls can therefore operate on the active workspace state without reinitializing browser context on every request.

\subsection{Workspace Read/Write Operations}
After initialization, the runtime follows a concrete execution path for each user prompt. The main steps are as follows:

\begin{enumerate}
    \item \textbf{Prompt-to-tool transition:} \\
    The user prompt is processed in the LangGraph loop. When the LLM determines that a model mutation is required, it emits a call to the bound \texttt{update\_html(new\_html)} tool.

    \item \textbf{Tool invocation and endpoint call:} \\
    The Python tool function \texttt{update\_html} sends a
    \[
    \texttt{POST /update-html}
    \]
    request to the Node automation bridge with the following payload:
    \[
    \{\texttt{"newHTML"}:\ \texttt{<generated-model-markup>}\}
    \]
    This endpoint is handled in \texttt{AI-Collaborator/puppeteer-server/index.js}, where the request is forwarded to \texttt{puppetEditModel(newHTML)}.

    \item \textbf{Browser-side model update:} \\
    Inside \texttt{puppetEditModel} (\texttt{AI-Collaborator/puppeteer-server/puppet.js}), Puppeteer calls a modeler-native update hook:
    \[
    \texttt{window.Modeler.updateModelViewHTML(newHTML)}
    \]
    This operation replaces the content of the active model instance through the modeler runtime API rather than direct DOM mutation.

    \item \textbf{Propagation to collaborators:} \\
    Because the mutation occurs directly in the live collaborative modeler DOM, the platform's existing synchronization channel propagates the resulting workspace change to connected participants. Thus, the agent update appears in the same shared session context as human edits.
\end{enumerate}

Additionally, read operations are available through \texttt{get\_current\_html()} (Python action) and \texttt{GET /get-html} (Node bridge), enabling the agent to fetch the latest workspace state before subsequent generation cycles. Validation is executed through \texttt{POST /validate-html}, which delegates to \texttt{window.Modeler.validateHTML(newHTML)} inside the Puppeteer session.

\section{Portability and Adaptation to Other Platforms}
While the implementation described in the preceding sections targets the Optimization Platform's modeler, the architecture was designed with explicit separation between \emph{platform-agnostic infrastructure} and \emph{platform-specific integration points}. This section identifies the concrete adaptation boundaries and explains how the same architecture can be ported to a different web application with minimal structural changes.

\subsection{Architectural Separation Principle}
The AI-collaboration pipeline is organized into three layers, each with a different scope of platform coupling:

\begin{enumerate}
    \item \textbf{Orchestration layer (Python/LangGraph):} Stateful workflow execution, LLM invocation, human-in-the-loop control, and conditional graph routing. This layer is fully platform-agnostic; it communicates with the environment exclusively through HTTP calls to the automation bridge. No platform-specific logic resides here except for the system prompt content.
    \item \textbf{Automation bridge (Node.js/Puppeteer):} Browser session management and HTTP endpoint exposure. The bridge is split into a generic session initialization component and platform-specific read/write/validate operations.
    \item \textbf{Frontend integration (chat runtime):} The assistant-ui chat runtime and its LangGraph adapter. The chat widget itself is framework-independent (React-based, embeddable); only the model-context injection logic is coupled to the host application's state representation.
\end{enumerate}

This layering means that porting the architecture to a different platform requires changes only at well-defined integration points, without restructuring the orchestration graph, the chat runtime protocol, or the bridge server skeleton.

\subsection{Automation Bridge: Generic vs.\ Platform-Specific Endpoints}
In the implemented bridge service (\texttt{AI-Collaborator/puppeteer-server/}), the code explicitly separates generic and platform-specific concerns. The session initialization endpoint (\texttt{POST /puppet}) and the Puppeteer bootstrap routine (\texttt{puppetInit}) are designed to work with any web application that has a login page. Authentication details are externalized as named constants:

\begin{itemize}
    \item \texttt{PLATFORM\_LOGIN\_URL}: the URL of the target application's login page.
    \item \texttt{PLATFORM\_EMAIL\_LOCATOR}, \texttt{PLATFORM\_PASSWORD\_LOCATOR},\\
    \texttt{PLATFORM\_SUBMIT\_BUTTON\_LOCATOR}: CSS selectors for the authentication form elements.
    \item \texttt{AI\_USER\_EMAIL}, \texttt{AI\_USER\_PASSWORD}: dedicated agent credentials.
\end{itemize}

To adapt session initialization to a different platform, a developer replaces these constants with values matching the target application's login page structure. The initialization flow itself (launch browser, navigate, authenticate, attach to workspace URL) remains unchanged.

The platform-specific endpoints (\texttt{POST /update-html}, \texttt{GET /get-html}, \texttt{POST /validate-html}) and their corresponding Puppeteer functions (\texttt{puppetEditModel}, \texttt{puppetGetModel}, \texttt{puppetValidateModel}) are explicitly documented as platform-dependent. These functions assume that the target application exposes certain methods on a global JavaScript object (in this case, \texttt{window.Modeler}) that the Puppeteer session can call via \texttt{page.evaluate()}. For a different platform, a developer would:

\begin{enumerate}
    \item Define equivalent read, write, and validate operations for the target application's DOM or API surface.
    \item Implement corresponding \texttt{page.evaluate()} calls that invoke the target application's exposed hooks.
    \item If the target application does not expose such hooks natively, implement thin adapter methods in the application's frontend code that bridge between the Puppeteer call and the internal application state.
\end{enumerate}

The bridge endpoint signatures (HTTP verbs, payload shapes, response formats) can remain identical, which means the orchestration layer requires no changes when the bridge internals are swapped.

\subsection{Orchestration Layer: Prompt as the Domain Interface}
The LangGraph workflow graph (node topology, conditional edges, interrupt/resume logic) is entirely domain-independent. The graph structure shown in the preceding sections -- generation, validation, repair, apply, and human-feedback states -- applies to any scenario where an LLM proposes structured changes to an editable workspace and a human reviews them before commitment.

The only component in the orchestration layer that carries platform-specific semantics is the \textbf{system prompt}. In the current implementation, the system prompt encodes the grammar rules, structural constraints, and output format expectations of the optimization modeler's markup language. For a different platform, the system prompt would be replaced with an equivalent specification of the target domain's representation language or editing conventions.

In practice, adapting the orchestration layer to a new domain requires:
\begin{itemize}
    \item writing a new system prompt that defines the target output format, structural rules, and domain role,
    \item and optionally adjusting the validation/repair prompts (\texttt{system\_prompt\_fix\_html}, \texttt{system\_prompt\_update\_html}) to reference the new domain's error semantics.
\end{itemize}

No changes to the graph topology, state schema, or routing logic are necessary.

\subsection{Frontend Integration: Embeddable Chat Runtime}
The assistant-ui chat widget communicates with the LangGraph backend through a standardized thread-and-run API contract (create, load, stream). This contract is not specific to any host application; it depends only on the LangGraph server endpoint and a graph identifier. Therefore, embedding the chat interface into a different web application requires:

\begin{enumerate}
    \item mounting the assistant-ui React component in the target application's UI container,
    \item configuring the LangGraph client URL and graph ID in the chat API module,
    \item and implementing a context-injection hook that retrieves the current application state and prepends it to each user message.
\end{enumerate}

The context-injection hook is the primary platform-specific adaptation point on the frontend side. In the current implementation, it queries the modeler's HTML representation; in a different application, it would query whatever state representation the target platform uses (e.g., a JSON document, a canvas state, or a code editor buffer).

\subsection{Summary of Adaptation Points}
\Cref{tab:adaptation-points} summarizes the concrete changes required to port the architecture to a different web platform.

\begin{table}[ht]
\centering
\caption{Platform-specific adaptation points for porting the AI collaborator architecture.}
\label{tab:adaptation-points}
\begin{tabular}{p{3.2cm} p{4.5cm} p{5.5cm}}
\toprule
\textbf{Layer} & \textbf{What to change} & \textbf{What remains unchanged} \\
\midrule
Automation bridge (init) & Login URL, form selectors, agent credentials & Browser launch, viewport setup, workspace-attach flow \\
\addlinespace
Automation bridge (ops) & \texttt{page.evaluate()} hooks for read/write/validate & HTTP endpoint signatures, bridge server skeleton \\
\addlinespace
Orchestration (Python) & System prompt, repair/update prompt templates & Graph topology, state schema, routing, HITL logic \\
\addlinespace
Frontend (chat) & Context-injection hook, UI mount point & assistant-ui runtime, LangGraph adapter, thread API \\
\bottomrule
\end{tabular}
\end{table}

Because the platform-specific surface is limited to configuration constants, a small set of browser-side hooks, and prompt content, the overall architecture can be reused across web applications without reimplementing the orchestration, chat runtime, or automation infrastructure.
